Equivalence Hypergraphs:\\ Semantics of E-Graphs and their Generalization to Monoidal Theories

### 2023-10-18 sections/combinatorial_semantics.tex

Before we begin we first fix some notation.
When $V$ is a set, we denote with $V^{*}$ a set of all finite sequences of the elements of $V$, i.e. $^*$ is a free monoid.
When $\phi : V \to V'$ is a function, we denote with $\phi^{*}$ an extension of $\phi$ to sequences, i.e. $\phi^{*} : V^* \to V'^{*}$ and it applies $\phi$ element-wise.
Similarly, in the case when $\psi \subset V \times V$ is a relation: we denote with $\psi^{*}: \subset V^* \times V^*$ the element-wise relation on ordered sequences.
We will denote with $A +_{f,g} B$ the pushout of $A \xleftarrow{f} C \xrightarrow{g} B$.

@@

We can not have $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1 < x_2$, i.e. the case where $x_1$ is not the immediate predecessor of $x_2$ because then it would mean that there exists at least $x_3$ such that $x_1 <_{\mu} x_2 <_{\mu} x_3$ and because $\phi$ should preserve them it would mean that $\phi(x_1) <_{\mu} \phi(x_3) <_{\mu} \phi(x_2)$ implying that $\phi(x_2)$ has two immediate predecessors: $\phi(x_1)$ and $\phi(x_3)$. Of course, if $\phi$ does not identify $x_1$ and $x_3$.

The first bullet is fine because it means that $\phi$ reflects the immediate predecessor. The second bullet leaves us with two options depicted below.

@@

    \[
    W \xrightarrow{f} X \xrightarrow{g} A \xleftarrow{h} Y \xleftarrow{k} Z
    \] 
    where $A \in obj(\catname{EHyp_{\Sigma}})$, $W, X, Y, Z$ are discrete and $f,g,h,k$ are monos.
    Since $f$ and $k$ are monos, $W$ and $Z$ are essentially subgraphs of $X$ and $Y$ and hence we will introduce the following notation which better depicts that such cospans are effectively e-hypergraphs with \textit{external} interfaces

@@

\Aleksei{Redraw this diagram in tikzit because importing 'quiver' introduces error}

@@

\[
X \xrightarrow{id_X} X \xrightarrow{id_X} X \xleftarrow{id_X} X \xleftarrow{id_X} X
\]

@@

since $X,Y,Z,Y', Y_{ext}$ are all discrete, $\uplus$ and $\setminus$ are standard set operations.

@@

Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into a SMC.
More precisely, below are cospans for tensor product and symmetry. Tensor unit is given by the unit in $\catname{EHyp_{\Sigma}}$.

% https://q.uiver.app/#q=WzAsMTcsWzAsMCwiZXh0KFgpIl0sWzEsMCwiWCJdLFsyLDAsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzAsMiwiZXh0KFgnKSJdLFsxLDIsIlgnIl0sWzIsMiwiQiJdLFszLDIsIlknIl0sWzQsMiwiZXh0KFknKSJdLFsyLDEsIlxcb3RpbWVzIl0sWzUsMSwiOj0iXSxbNiwxLCJleHQoWCkrZXh0KFgnKSJdLFs4LDEsIlgrWCciXSxbMTAsMSwiQStCIl0sWzEyLDEsIlkrWSciXSxbMTQsMSwiZXh0KFkpK2V4dChZJykiXSxbMCwxLCJmX3tleHR9Il0sWzQsMywiZ197ZXh0fSIsMl0sWzMsMiwiZyIsMl0sWzEsMiwiZiJdLFs1LDYsImYnX3tleHR9Il0sWzYsNywiZiciXSxbOCw3LCJnJyIsMl0sWzksOCwiZydfe2V4dH0iLDJdLFsxMiwxMywiZl97ZXh0fStmJ197ZXh0fSJdLFsxMywxNCwiZitmJyJdLFsxNSwxNCwiZytnJyIsMl0sWzE2LDE1LCJnX3tleHR9ICsgZydfe2V4dH0iLDJdXQ==
\[
\adjustbox{scale=0.6, center}{
\begin{tikzcd}
	{X_{ext}} & X & A & Y & {Y_{ext}} \\
	&& \otimes &&& {:=} & {X_{ext}+ext(X')} && {X+X'} && {A+B} && {Y+Y'} && {Y_{ext}+Y'_{ext}} \\
	{ext(X')} & {X'} & B & {Y'} & {Y'_{ext}}
	\arrow["{f_{ext}}", from=1-1, to=1-2]
	\arrow["{g_{ext}}"', from=1-5, to=1-4]
	\arrow["g"', from=1-4, to=1-3]
	\arrow["f", from=1-2, to=1-3]
	\arrow["{f'_{ext}}", from=3-1, to=3-2]
	\arrow["{f'}", from=3-2, to=3-3]
	\arrow["{g'}"', from=3-4, to=3-3]
	\arrow["{g'_{ext}}"', from=3-5, to=3-4]
	\arrow["{f_{ext}+f'_{ext}}", from=2-7, to=2-9]
	\arrow["{f+f'}", from=2-9, to=2-11]
	\arrow["{g+g'}"', from=2-13, to=2-11]
	\arrow["{g_{ext} + g'_{ext}}"', from=2-15, to=2-13]
\end{tikzcd}
}
\]

% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
\[
\adjustbox{scale=0.75, center}{
\begin{tikzcd}
	{X + Y} && {X + Y} && {Y + X} && {Y + X} && {Y + X}
	\arrow["{id_X + id_Y}", from=1-1, to=1-3]
	\arrow["{\sigma_{X,Y}}", from=1-3, to=1-5]
	\arrow["{id_Y + id_X}"', from=1-9, to=1-7]
	\arrow["{id_Y + id_X}"', from=1-7, to=1-5]
\end{tikzcd}
}
\]

@@

\begin{remark}
    Pushout for composition is computed by first taking a coproduct of $A$ and $B$ and then identifying the nodes that share the same pre-image in $Y$.
    Consider an example in Figure~\ref{fig:e-hyp_composition_example} below.
    The pushout is depicted in Figure~\ref{fig:e-hyp_composition_example_2}.
    New interfaces are shown in Figure~\ref{fig:e-hyp_composition_example_3}.
    They are computed as shown above: a disjoint union of the input interface of the left component of the composition with the input interface of the right component of the composition except for the nodes in the image of the outermost output interface of the left component is taken.
    Similarly for the output interface.
    In particular, $\{0,1,2,3\} \uplus (\{6,7,8,9,10\} \setminus \{6\}) = \{0,1,2,3,7,8,9,10\}$.
    The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
    \update{We should probably require that the image of $f_{ext};f{}$ has no predecessors and has no incoming edges}
    Note that pushout for such composition always exists as the nodes that are being identified are unrelated: they are not in conflict with anything and the sets of their predecessors are empty.
\end{remark}

@@

\Aleksei{what is $h$? $k$? $f_ext: ext(X) -> X$, so how is $f_ext$ also of type $ext(X) -> ext(X) + X + X'$ ? }
\begin{definition}
    $+$ of two morphisms 
    \[
    X_{ext} \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} Y_{ext}
    \]
    \[
    X_{ext} \xrightarrow{f'_{ext}} X' \xrightarrow{f'} B \xleftarrow{g'} Y' \xleftarrow{g'_{ext}} Y_{ext}
    \]

@@

    \[
    X_{ext} \xrightarrow{f_{ext}} X_{ext} + X + X' \xrightarrow{h + f+f'} C \xleftarrow{k + g+g'} Y_{ext} + Y + Y' \xleftarrow{g_{ext}} Y_{ext}
    \]
    \label{morphism:composition_ecospans}

    $A$ and $B$ are generic e-hypergraphs that represent the carriers of the morphisms.
    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now in conflict with all nodes and edges of $B$ and with a new hierarchical edge that has $|X_{ext}|$ sources and $|Y_{ext}|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ in $A$ and $B$ such that $[e) = \varnothing$ and $[v)$. And such that its sources and targets are top-level nodes. $h$ and $k$ are morphisms that map $X_{ext}$ and $Y_{ext}$ to these sources and targets respectively. Since both $A$ and $B$ have the same outermost input and output interfaces they are added as the outermost interfaces of their $C$.
\end{definition}

@@

\begin{enumerate}
        \item $A$ is a directed acyclic e-hypergraph.
        \item $f_{ext}$, $f$, $g$, $g_{ext}$ are monomorphic.
        \item The in-degree and out-degree of every node in $A$ are at most 1.
        \item Nodes with the in(out)-degree of 0 are exactly the nodes in the image of $f$ (respectively, $g$).
        \item $[v) \not = \varnothing$ for all $v$ in the image of $f$ ($g$) which pre-image is not in $X_{ext}$ ($Y_{ext}$). 
        \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
\end{enumerate}
\label{def:monogamy_ehyp}

@@

\begin{lemma}
    \begin{enumerate}
        \item The tensor product of two mda-cospans is an mda-cospan.
        \item $+$ of two mda-cospans is an mda-cospan.
    \end{enumerate}
    \begin{proof}
        Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
        Hence, everything is preserved component-wise by copairing and coproduct.
    \end{proof}
    \begin{proof}
        By construction in Figure~\ref{fig:A+B}.
    \end{proof}
\end{lemma}

% \begin{proposition}
%     Id morphism

%     \[
%     X \xrightarrow{id_{X}} X \xrightarrow{id_{X}} X \xleftarrow{id_{X}} X \xleftarrow{id_X} X
%     \]

%     is an mda-cospan.
%     \begin{proof}
%         Proof by checking all the constraints above. Discrete e-hypergraphs is obviously directed and acyclic. Id-morphisms are monos. The degree of all the nodes is 0 and all the nodes are in the image of $id_{X}$. Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.
%     \end{proof}
% \end{proposition}

\update{Anything about symmetry?}

@@

    Narrowing the morphisms of $\Ecospans$ to mda-cospans gives us a subcategory $\MdaEcospans$.
    Id morphims is given by the following cospan

    \[
    X \xrightarrow{id_{X}} X \xrightarrow{id_{X}} X \xleftarrow{id_{X}} X \xleftarrow{id_X} X
    \]

    which is an mda-cospan by checking all the necessary constraints in~\ref{def:monogamy_ehyp}.
    Discrete e-hypergraph $X$ is obviously directed and acyclic.
    Id-morphisms are monos.
    The degree of all the nodes is 0 and all the nodes are in the image of $id_{X}$.
    Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.

    Composition is given in~\ref{morphism:composition_ecospans} and composing two mda-cospans yields an mda-cospan.
    Pushout glues out-degree 0 nodes on the left and in-degree 0 nodes on the right which does not introduce any cycles.
    This also does not increase their degree to more than 1.
    $f_{ext}$ is mono because its image $X$ is unchanged.
    The image of $B \xrightarrow{i_2} A + B$ is mono, $g$ is mono and their composition is hence mono.
    Pushout is then obtained by identifying the nodes with the pre-image in $Y_{ext}$ which are removed in $Y' \setminus Y_{ext}$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus Y_{ext}$. $g';i_2$ yields $h$ and therefore $h$ is mono.
    Likewise, $k$ is mono.

@@

% show that EHyp is symmetric monoidal

@@

\begin{definition}[Well-typed edge]

    Consider a hierarchical edge $e$ in $E_{A}$ and the sets $V_{e, in}$ and $V_{e,out}$ of its input and output nodes respectively. Then consider partitions $V_{e, in}^{\bar{\#}}$ and $V_{e,out}^{\bar{\#}}$  of the nodes according to $\bar{\#}$. We call $e$ well-typed if for all $v \in V_{e,in}^{\bar{\#}}$ $|v| = |s(e)|$ and for all $v \in V_{e,out}^{\bar{\#}}$ $|v| = |t(e)|$. We will call an mda-cospan well-typed if all hierarchical edges in its carrier are well-typed.
\end{definition}

@@

\begin{proposition}
    Id-cospan

    \[
    X \xrightarrow{id_X} X \xrightarrow{id_X} X \xleftarrow{id_X} X \xleftarrow{id_X} X
    \]

    is well-typed.
    \begin{proof}
        By definition, we only include well-typed morphisms.
    \end{proof}
\end{proposition}

\begin{proposition}[The tensor product of two mda-cospans is again an mda-cospan]
    The tensor product does not introduce any new hierarchical edges and images of the interfaces are
preserved by copairing.
\end{proposition}

\begin{proposition}[Composition of two well-typed mda-cospans is a well-typed mda-cospan]
     Composition does not introduce any new hierarchical edges.
Furthermore, as composition identifies only nodes in the image of the outermost interfaces, input and
output nodes of hierarchical edges are not affected.
\end{proposition}

@@

The propositions above make $\WellTypedMdaEcospans$ into a proper symmetric monoidal category. There is an obvious inclusion

@@

\begin{definition}[\cite{Frobenius}]
A rewriting system $\mathcal{R}$ in a PROP $\mathbb{A}$ consists of a set of rewriting rules, i.e. pairs $\langle l , r \rangle$ of morphisms $l , r : i \to j$ in $\mathbb{A}$ with the same arities and co-arities. Given $a, b : m \to n$ in $\mathbb{A}$, $a$ rewrites into $b$ via $\mathcal{R}$, written $a\Rightarrow_{\mathcal{R}} b$ ,if they are decomposable as follows, for some rule $\langle l,r \rangle \in \mathcal{R}$    
\end{definition}

@@

    \[
    \scalebox{0.75}{\tikzfig{combinatorial_semantics/prop_rewriting}}
    \]
\end{figure}

@@

If they are decomposable as above then we say that $a$ rewrites to $b$ under rule $\langle l, r \rangle$ in $\mathcal{R}^{+}$. Another case is when some $a'$ rewrites to some $b'$ under rule $\langle l, r \rangle$ in the sense above and then we say that $a$ rewrites to $b$ under the same rule if they are decomposable as below.

@@

    \[
    \scalebox{0.6}{
    \tikzfig{combinatorial_semantics/prop_plus_rewriting}
    }
    \]
\end{figure}

@@

\begin{theorem}[Theorem 35~\cite{Frobenius2}]
    Let $R$ be any rewriting system on $\textsf{PROP}(\Sigma)$. Then,
    \[
    g \Rightarrow_{\mathcal{R}} h \;\text{ iff }\; \llbracket ^{\ulcorner}g^{\urcorner} \rrbracket \Rrightarrow_{\llbracket \mathcal{R} \rrbracket} \llbracket ^{\ulcorner}h^{\urcorner} \rrbracket 
    \]

    where $\llbracket \cdot \rrbracket : \textsf{PROP}(\Sigma) \to \MdaCospans$ is a faithful monoidal functor.
\end{theorem}

\begin{remark}
    Note that even though $n \xrightarrow{} G \xleftarrow{} m$ and $0 \xrightarrow{} G \xleftarrow{} n + m$ are essentially isomorphic \update{(they are not technically isomorphic because they have different legs)}, the functor $\llbracket \cdot \rrbracket$ deals with arbitrary string diagrams (that have multiple inputs and multiple outputs) in $\textsf{PROP}(\Sigma)$ and interprets them as cospans $n \xrightarrow{} G \xleftarrow{} m$. However, there is a syntactic operation $\ulcorner \cdot \urcorner$ on $\textsf{PROP}(\Sigma)$ to translate between string diagrams with multiple inputs and outputs and string diagrams with only outputs.

    \[
\tikzfig{combinatorial_semantics/bending_wires}
    \]

>>>>

Before we begin we first fix some notation.
When $V$ is a set, we denote with $V^{*}$ a set of all finite sequences of the elements of $V$, i.e. $^*$ is a free monoid.
When $\phi : V \to V'$ is a function, we denote with $\phi^{*}$ an extension of $\phi$ to sequences, i.e. $\phi^{*} : V^* \to V'^{*}$ and it applies $\phi$ element-wise.
Similarly, in the case when $\psi \subset V \times V$ is a relation: we denote with $\psi^{*}: \subset V^* \times V^*$ the element-wise relation on ordered sequences.
We will denote with $A +_{f,g} B$ the pushout of $A \xleftarrow{f} C \xrightarrow{g} B$ and will use $i_j$ to denote the $j^{\text{th}}$ injection into a coproduct of the form $X_{j} \xrightarrow{i_{j}} X_{1} + \ldots X_{j} + \ldots X_{n}$ where $+$ denotes the coproduct.

@@

We can not have $\phi(x_1) <_{\mu} \phi(x_2)$ and $x_1 < x_2$, i.e. the case where $x_1$ is not the immediate predecessor of $x_2$ because then it would mean that there exists at least $x_3$ such that $x_1 <_{\mu} x_3 <_{\mu} x_2$ and because $\phi$ should preserve them it would mean that $\phi(x_1) <_{\mu} \phi(x_3) <_{\mu} \phi(x_2)$ implying that $\phi(x_2)$ has two immediate predecessors: $\phi(x_1)$ and $\phi(x_3)$.
Of course, if $\phi$ does not identify $x_1$ and $x_3$.
The first bullet is fine because it means that $\phi$ reflects the immediate predecessor.
The second bullet leaves us with two options depicted above.

@@

    \[
    W \xrightarrow{f} X \xrightarrow{g} A \xleftarrow{h} Y \xleftarrow{k} Z
    \]

    where
    \begin{enumerate}
        \item $A \in \text{obj}(\catname{EHyp_{\Sigma}})$
        \item $W, X, Y, Z$ are discrete and $f,g,h,k$ are monos
        \item \update{The image of $f;g$ (respectively, $k;h$) constitutes of nodes $v$ such that $[v) = \varnothing$}
    \end{enumerate}
    Since $f$ and $k$ are monos, $W$ and $Z$ are essentially subgraphs of $X$ and $Y$ and hence we will introduce the following notation which better depicts that such cospans are effectively e-hypergraphs with \textit{external} interfaces

@@

@@

\[
X \xrightarrow{id_X} X \xrightarrow{id_X} X \xleftarrow{id_X} X \xleftarrow{id_X} X
\]\label{morphism:identity}

@@

since $X,Y,Z,Y', Y_{ext}$ are all discrete, $\uplus$ and $\setminus$ are standard set operations.
$h = a;p_1 + g|_{Y' \setminus Y_{ext}};p_2$, $k = b;p_2 + f|_{Y \setminus Y_{ext}};p_1$.
\question{Is this a standard syntax for a narrowed function?}

@@

Similarly to~\ref{def:cspd} this category has a symmetric monoidal product given by a coproduct in $\catname{EHyp_{\Sigma}}$ which makes it into an SMC.
More precisely, below are cospans for tensor product and symmetry. Tensor unit is given by the unit in $\catname{EHyp_{\Sigma}}$.

% https://q.uiver.app/#q=WzAsMTcsWzAsMCwiZXh0KFgpIl0sWzEsMCwiWCJdLFsyLDAsIkEiXSxbMywwLCJZIl0sWzQsMCwiZXh0KFkpIl0sWzAsMiwiZXh0KFgnKSJdLFsxLDIsIlgnIl0sWzIsMiwiQiJdLFszLDIsIlknIl0sWzQsMiwiZXh0KFknKSJdLFsyLDEsIlxcb3RpbWVzIl0sWzUsMSwiOj0iXSxbNiwxLCJleHQoWCkrZXh0KFgnKSJdLFs4LDEsIlgrWCciXSxbMTAsMSwiQStCIl0sWzEyLDEsIlkrWSciXSxbMTQsMSwiZXh0KFkpK2V4dChZJykiXSxbMCwxLCJmX3tleHR9Il0sWzQsMywiZ197ZXh0fSIsMl0sWzMsMiwiZyIsMl0sWzEsMiwiZiJdLFs1LDYsImYnX3tleHR9Il0sWzYsNywiZiciXSxbOCw3LCJnJyIsMl0sWzksOCwiZydfe2V4dH0iLDJdLFsxMiwxMywiZl97ZXh0fStmJ197ZXh0fSJdLFsxMywxNCwiZitmJyJdLFsxNSwxNCwiZytnJyIsMl0sWzE2LDE1LCJnX3tleHR9ICsgZydfe2V4dH0iLDJdXQ==
\[
\adjustbox{scale=0.6, center}{
\begin{tikzcd}
	{X_{ext}} & X & A & Y & {Y_{ext}} \\
	&& \otimes &&& {:=} & {X_{ext}+ext(X')} && {X+X'} && {A+B} && {Y+Y'} && {Y_{ext}+Y'_{ext}} \\
	{ext(X')} & {X'} & B & {Y'} & {Y'_{ext}}
	\arrow["{f_{ext}}", from=1-1, to=1-2]
	\arrow["{g_{ext}}"', from=1-5, to=1-4]
	\arrow["g"', from=1-4, to=1-3]
	\arrow["f", from=1-2, to=1-3]
	\arrow["{f'_{ext}}", from=3-1, to=3-2]
	\arrow["{f'}", from=3-2, to=3-3]
	\arrow["{g'}"', from=3-4, to=3-3]
	\arrow["{g'_{ext}}"', from=3-5, to=3-4]
	\arrow["{f_{ext}+f'_{ext}}", from=2-7, to=2-9]
	\arrow["{f+f'}", from=2-9, to=2-11]
	\arrow["{g+g'}"', from=2-13, to=2-11]
	\arrow["{g_{ext} + g'_{ext}}"', from=2-15, to=2-13]
\end{tikzcd}
}
\]\label{morphism:tensoring}

% https://q.uiver.app/#q=WzAsNSxbMCwwLCJYICsgWSJdLFsyLDAsIlggKyBZIl0sWzQsMCwiWSArIFgiXSxbNiwwLCJZICsgWCJdLFs4LDAsIlkgKyBYIl0sWzAsMSwiaWRfWCArIGlkX1kiXSxbMSwyLCJcXHNpZ21hX3tYLFl9Il0sWzQsMywiaWRfWSArIGlkX1giLDJdLFszLDIsImlkX1kgKyBpZF9YIiwyXV0=
\[
\adjustbox{scale=0.75, center}{
\begin{tikzcd}
	{X + Y} && {X + Y} && {Y + X} && {Y + X} && {Y + X}
	\arrow["{id_X + id_Y}", from=1-1, to=1-3]
	\arrow["{\sigma_{X,Y}}", from=1-3, to=1-5]
	\arrow["{id_Y + id_X}"', from=1-9, to=1-7]
	\arrow["{id_Y + id_X}"', from=1-7, to=1-5]
\end{tikzcd}
}
\]\label{morphism:symmetry}

@@

\begin{remark}
    Pushout for composition is computed by first taking a coproduct of $A$ and $B$ and then identifying the nodes that share the same pre-image in $Y_{ext}$.
    Consider an example in Figure~\ref{fig:e-hyp_composition_example} below.
    The pushout is depicted in Figure~\ref{fig:e-hyp_composition_example_2}.
    New interfaces are shown in Figure~\ref{fig:e-hyp_composition_example_3}.
    They are computed as shown above: a disjoint union of the input interface of the left component of the composition with the input interface of the right component of the composition except for the nodes in the image of the outermost output interface of the left component is taken.
    Similarly for the output interface.
    In particular, $\{0,1,2,3\} \uplus (\{6,7,8,9,10\} \setminus \{6\}) = \{0,1,2,3,7,8,9,10\}$.
    The conflict and child relations in the pushout are a disjoint union of the corresponding relations from pushout components.
    Note that pushout for such composition always exists as the nodes that are being identified are unrelated: they are not in conflict with anything and the sets of their predecessors are empty.
\end{remark}

@@

\begin{definition}
    $+$ of two morphisms 
    \[
    X_{ext} \xrightarrow{f_{ext}} X \xrightarrow{f} A \xleftarrow{g} Y \xleftarrow{g_{ext}} Y_{ext}
    \]
    \[
    X_{ext} \xrightarrow{f'_{ext}} X' \xrightarrow{f'} B \xleftarrow{g'} Y' \xleftarrow{g'_{ext}} Y_{ext}
    \]

@@

    \[
    X_{ext} \xrightarrow{i_{1}} X_{ext} + X + X' \xrightarrow{h + f+f'} C \xleftarrow{k + g+g'} Y_{ext} + Y + Y' \xleftarrow{i_1} Y_{ext}
    \]
    \label{morphism:composition_ecospans}

    $A$ and $B$ are generic e-hypergraphs that represent the carriers of the morphisms.
    $C$ is a coproduct of e-hypergraphs $A$ and $B$ with the extra requirement that all nodes and edges of $A$ are now in conflict with all nodes and edges of $B$ and with a new hierarchical edge that has $|X_{ext}|$ sources and $|Y_{ext}|$ targets and is the immediate predecessor of the edges $e$ and nodes $v$ in $A$ and $B$ such that $[e) = \varnothing$ and $[v) = \varnothing$.
    And such that its sources and targets are top-level nodes.
    $h$ and $k$ are morphisms that map $X_{ext}$ and $Y_{ext}$ to these sources and targets respectively. In particular, in Figure~\ref{fig:A+B}, $h = \{c_1 \mapsto c_1, \ldots, c_n \mapsto c_n\}$ and $k = \{d_1 \mapsto d_1, \ldots, d_l \mapsto d_l\}$.
    Since both $A$ and $B$ have the same outermost input and output interfaces they are added as the outermost interfaces of their $C$.
\end{definition}

\begin{remark}
    We can generalise the above definition to a plus of $n$ cospans in an obvious way.
\end{remark}

@@

\begin{enumerate}
        \item $A$ is a directed acyclic e-hypergraph.
        \item $f_{ext}$, $f$, $g$, $g_{ext}$ are monomorphic.
        \item The in-degree and out-degree of every node in $A$ are at most 1.
        \item Nodes with the in(out)-degree of 0 are exactly the nodes in the image of $f$ (respectively, $g$).
        \item $[v) \not = \varnothing$ for all $v$ in the image of $f$ (respectively, $g$) which pre-image is not in $X_{ext}$ (respectively, $Y_{ext}$). 
        % the item below is not assumed by the definition of a cospan of e-hypergraphs
        % \item $[v) = \varnothing$ for all $v$ in the image of $f_{ext};f$, and for all $v$ in the image of $g_{ext};g$. 
\end{enumerate}
\label{def:monogamy_ehyp}

@@

% \begin{lemma}
%     \begin{enumerate}
%         \item The tensor product of two mda-cospans is an mda-cospan.
%         \item $+$ of two mda-cospans is an mda-cospan.
%     \end{enumerate}
%     \begin{proof}
%         Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
%         Hence, everything is preserved component-wise by copairing and coproduct.
%     \end{proof}
%     \begin{proof}
%         By construction in Figure~\ref{fig:A+B}.
%     \end{proof}
% \end{lemma}

@@

    Narrowing the morphisms of $\Ecospans$ to mda-cospans gives us a subcategory $\MdaEcospans$.

@@

\begin{proposition}
    $\MdaEcospans$ is an SMC.
\begin{proof}
To show the statement above we need to first show that $\MdaEcospans$ is a proper category and then show that it has a symmetry, tensor product and the tensor unit.
In particular we need to show that the cospans that correspond to particular morphisms are mda-ones according to~\ref{def:monogamy_ehyp}.
\begin{enumerate}
    \item Id cospan is an mda-cospan. Recall the id cospan in~\ref{morphism:identity}.
    Discrete e-hypergraph $X$ is obviously directed and acyclic.
    Id-morphisms are monos.
    The degree of all the nodes is 0 and all the nodes are in the image of $id_{X}$.
    Set of nodes for bullet (5) is empty and $[v) = \varnothing$ for bullet (6) is implied by discreteness.
    \item Composition~\ref{morphism:composition_ecospans} of two mda-cospans yields an mda-cospan.
    Pushout glues out-degree 0 nodes on the left and in-degree 0 nodes on the right which does not introduce any cycles.
    This also does not increase their degrees to more than 1.
    $f_{ext}$ (respectively, $g_{ext}$) is mono because its image $X$ (respectively, $Z$) is unchanged.
    The morphism $B \xrightarrow{i_2} A + B$ is mono, $g$ is mono and their composition is hence mono.
    Also note that $p_2$ narrowed to $B \setminus Y_{ext}$ is exactly $i_2$ and similarly for $p_1$ and $i_1$.
    Pushout is then obtained by identifying the nodes in $A + B$ that share their pre-image in $Y_{ext}$.
    These nodes are removed in $Y' \setminus Y_{ext}$ and hence $g';i_2$ is mono where $g'$ is $g$ narrowed to $Y' \setminus Y_{ext}$.
    $g';i_2$ is equal to $g';p_2$ because the image of $g'$ does not contain any nodes with pre-image in $Y_{ext}$.
    Next, $a$ is mono and $a;p_1$ is mono, because $p_1$ only identifies nodes with pre-image in $Y_{ext}$.
    $h = a;p_1 + g';p_2$ and hence $h$ is mono.
    Likewise, $k$ is mono.
    \item Tensor product~\ref{morphism:tensoring} of two mda-cospans is an mda-cospan.
          Tensoring does not introduce any new edges or nodes apart from those already in $A$ and $B$.
          Hence, everything is preserved component-wise by copairing and coproduct.
    \item Symmetry~\ref{morphism:symmetry} is an mda-cospan.
          $X$ and $Y$ are both discrete as well as their coproduct and $\sigma_{X,Y}$ is obviously mono.
          Furthermore, all nodes of $X + Y$ are in the image of $\sigma_{X,Y}$ and $id_{X} + id_{Y}$.
    \item The tensor unit is discrete because it is an empty e-hypergraph and hence does not break monogamous property.
\end{enumerate}

The bullets above makes $\MdaEcospans$ into a proper SMC.
\end{proof}    
\end{proposition}

\question{Proofs ok?}

\begin{proposition}
    $+$ of two mda-cospans is an mda-cospan.
    \begin{proof}
        By construction in Figure~\ref{fig:A+B}.
    \end{proof}
\end{proposition}

@@

\begin{definition}[Well-typed cospan]

    Consider a hierarchical edge $e$ in $E_{A}$ and the sets $V_{e, in}$ and $V_{e,out}$ of its input and output nodes respectively.
    Then consider partitions $V_{e, in}^{\bar{\#}}$ and $V_{e,out}^{\bar{\#}}$  of the nodes according to $\bar{\#}$.
    We call $e$ well-typed if for all $v \in V_{e,in}^{\bar{\#}}$ $|v| = |s(e)|$ and for all $v \in V_{e,out}^{\bar{\#}}$ $|v| = |t(e)|$.
    We will call an mda-cospan in $\MdaEcospans$ well-typed if all hierarchical edges in its carrier are well-typed.
\end{definition}

@@

\begin{proposition}
    $\WellTypedMdaEcospans$ is a proper SMC.
\begin{proof}
    As before, we need to show the well-typedness of the key ingredients of an SMC.
    \begin{enumerate}
        \item Id-cospan~\ref{morphism:identity} is well-typed.
              Discrete e-hypergraphs do not have any hierarchical edges and hence are well-typed.
        \item Composition~\ref{morphism:composition_ecospans} of two well-typed mda-cospans is a well-typed mda-cospan.
              Composition does not introduce any new hierarchical edges.
              Furthermore, as composition identifies only nodes in the image of the outermost interfaces, input and
              output nodes of hierarchical edges are not affected and the number of sources and targets of an edge remains unchanged as well.
        \item The tensor product of two mda-cospans is again an mda-cospan.
              The tensor product does not introduce any new hierarchical edges and does not affect the existing hierarchical edges.
        \item The tensor unit is descrete and hence tensoring with it does not break well-typedness.
    \end{enumerate}
    All the items above show that $\WellTypedMdaEcospans$ is indeed an SMC.
\end{proof}
\end{proposition}

@@

There is an obvious inclusion

@@

\begin{definition}[\cite{Frobenius}]
A rewriting system $\mathcal{R}$ in a PROP $\mathbb{A}$ consists of a set of rewriting rules, i.e. pairs $\langle l , r \rangle$ of morphisms $l , r : i \to j$ in $\mathbb{A}$ with the same arities and co-arities.
Given $a, b : m \to n$ in $\mathbb{A}$, $a$ rewrites into $b$ via $\mathcal{R}$, written $a\Rightarrow_{\mathcal{R}} b$ ,if they are decomposable as in Figure~\ref{fig:prop_rewriting}, for some rule $\langle l,r \rangle \in \mathcal{R}$    
\end{definition}

@@

    \[
    \scalebox{0.75}{\tikzfig{combinatorial_semantics/prop_rewriting}}
    \]
    \caption{Rewriting in $\textsf{PROP}(\Sigma)$}
    \label{fig:prop_rewriting}
\end{figure}

@@

If they are decomposable as in Figure~\ref{fig:prop_rewriting} then we say that $a$ rewrites to $b$ under rule $\langle l, r \rangle$ in $\mathcal{R}^{+}$.
Another case is when some $a'$ rewrites to some $b'$ under rule $\langle l, r \rangle$ in the sense above ($\Rightarrow_{\mathcal{R}})$)) and then we say that $a$ rewrites to $b$ under the same rule if they are decomposable as in Figure~\ref{fig:enriched_prop_rewriting}.

@@

    \[
    \scalebox{0.6}{
    \tikzfig{combinatorial_semantics/prop_plus_rewriting}
    }
    \]
    \caption{Rewriting in $\textsf{PROP}^{+}(\Sigma)$}
    \label{fig:enriched_prop_rewriting}
\end{figure}

@@

\textbf{Update: 4 Oct}

\begin{theorem}[Theorem 35~\cite{Frobenius2}]
\label{theorem:rewriting_no_frob}
    Let $R$ be any rewriting system on $\textsf{PROP}(\Sigma)$. Then,
    \[
    g \Rightarrow_{\mathcal{R}} h \;\text{ iff }\; \llangle ^{\ulcorner}g^{\urcorner} \rrangle \Rrightarrow_{\llangle \mathcal{R} \rrangle} \llangle ^{\ulcorner}h^{\urcorner} \rrangle 
    \]

    where $\llangle \cdot \rrangle : \textsf{PROP}(\Sigma) + Frob \to \catname{Csp_{D}(Hyp_{\Sigma})}$ is an isomorphism of props.
    \update{
        Last time I confused functors $\llangle \cdot \rrangle$ and $\llbracket \cdot \rrbracket$.
        Note that the result above is a weaker version of the same result for rewriting in $\textsf{PROP}(\Sigma) + Frob$.
        $Frob$ is an SMT with the signature $\Sigma_{Frob}$ which constitutes of \textit{only} Frobenius generators and terms are quotiented by both SMC laws and Frobenius equations.
        So, essentially, $Frob = \textsf{PROP}(\Sigma_{Frob}, \mathcal{E}_{Frob})$. 
        \question{Is it true that by quotienting $\textsf{PROP}(\Sigma)$ by SMC laws and quotienting $Frob$ by SMC and Frob laws we automatically quotient $\textsf{PROP}(\Sigma) + Frob$ by SMC and Frob laws?} 
    }
\end{theorem}

\begin{definition}[$\ulcorner \cdot \urcorner$]
